### 2016-11-07 ichol_writeup.tex

\documentclass[12pt,letterpaper]{article}

%\usepackage{cvpr}
%\documentclass[11pt,letterpaper]{article}
\setlength{\textheight}{8.875in}
\setlength{\textwidth}{6.875in}
\setlength{\columnsep}{0.3125in}
\setlength{\topmargin}{0in}
\setlength{\headheight}{0in}
\setlength{\headsep}{0in}
\setlength{\parindent}{1pc}
\setlength{\oddsidemargin}{-.304in}
\setlength{\evensidemargin}{-.304in}

\usepackage{times}
\usepackage{bbm}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{caption}
\usepackage{subcaption}
\DeclareMathOperator*{\argmin}{\arg\!\min}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{coro}{Corollary}
\newtheorem{prop}{Property}
\newtheorem{definition}{Definition}

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).

\def\texte#1{#1}

\begin{document}
\def\RR{\mathbb{R}}
\def\PP{\mathbb{P}}
\def\AA{\mathbb{A}}
\def\LL{\mathbb{L}}
\def\SS{\mathbb{S}}
\def\barr{\bar{\mathbb{R}}}
\def\mat#1{{\mathcal{#1}}}
\def\vect#1{\mbox{\boldmath $#1$}}
\def\PPi{\mbox{\boldmath$\Pi$}}
\def\squig{\rightsquigarrow}
%\def\vect#1{#1}
%\def\mat#1{#1}
\def\comment#1{{}}
%\def\qmatrix#1{\left[\begin{array}{l}#1\end{array}\right]}
\def\qmatrix#1{\left[\begin{matrix}#1\end{matrix}\right]}

\title{Kernel Square-Loss Examplar Machines}
\author{Rafael Rezende}
%\date{}
\maketitle

\section{Incomplete Cholesky decomposition}

In previous write-ups, we wanted to extend a $n\times r$ low-rank decomposition of a $n\times n$ kernel matrix $K$ into a $(n+1)\times (r+1)$ of the matrix 
\begin{equation}
K'=\begin{pmatrix}
K & k_0 \\
k_0^T & k_{00}
\end{pmatrix}.
\end{equation}

We assumed the $(n+1)$-th column of this matrix decomposition $B'$ would be all zero except for the last writing
\begin{equation}
B'=\begin{pmatrix}
B & 0 \\
v^T & u
\end{pmatrix},\label{eq:old}
\end{equation}
where $B$ is the incomplete Cholesky decomposition (ICD) of $K$.
However, the pivot algorithm for ICD states $B'$ should be written as
\begin{equation}
B'=\begin{pmatrix}
B & w \\
v^T & u
\end{pmatrix},\label{eq:new}
\end{equation}
with $u$ in $\RR$, $v$ in $\RR^r$ and $w$ in $\RR^n$. 
Decomposing $K'\approx B'B'^T$ gives us 
\begin{align}
\begin{cases}
Bv+uw=k_0,\\
u+||v||^2=k_{00}.\label{syst:comp}
\end{cases}
\end{align}

Let $I_n=\{i_1,i_2,...,i_r\}\subset \{1,2,...,n\}$ be the set of pivots of $B$ and $J_n$ be the sorted complement if $I_n$. 
We add one more sample and we wish to decompose $K'$ in a $(n+1)\times (r+1)$ ICS. 
For this, we add another pivot, making $I_{n+1} = I_n\cup \{n+1\}$ the set of pivots of $K'$. Following the ICS algorithm gives us a solution to Equation (\ref{syst:comp}) in the form of 
\begin{align}
\begin{cases}
\vspace{3 mm}
w(I_n) = 0, v = B(I_n,:)^+k_0(I_n),\\
\vspace{3 mm}
u = \sqrt{k_{00}-||v||^2},\\
w(J_n) = \dfrac{1}{u}\left(K_0(J_n)-B(J_n,:)v \right),
\end{cases}
\end{align}\label{syst:sol}
where $B(I,:)$ is the block matrix of $B$ with rows $I$ and all columns.\\
Equation (\ref{syst:sol}) follows the same solution we had for (\ref{eq:old}), except by the third line. $B'$ from Equations (\ref{eq:old}) and (\ref{eq:new}) do not give same approximation of $K'$.

\section{Real data}
With the ICS algorithm applied to real data ($4096$ dimensional normalized CNN features from Holidays dataset) for a big enough sample of data $n$, we consistently find $r=4096$, which is smaller than $n$ but not low dimensional. 
Also, when adding the $(r+1)$-th dimension of $B'$, we find that $||v||=k_{00}$ and $w$ is non defined from Equation (\ref{syst:sol}).

\section{Eliminating negatives from negative encoding}

\end{document}