### 2016-11-07 sec_kernel_methods.tex

\section{The kernel SLEM}
\label{nonlinear SLEM}

\subsection{Introduction to kernel methods}\label{kernel:review}
Let us recall a few basic facts about kernel methods for supervised
classification. We consider a reproducing kernel Hilbert space (RKHS)
$H$ formed by real functions over some set
$X$, and denote by $k$ the corresponding reproducing kernel.  We
address the following learning problem over $H\times\RR$:
\begin{equation}
\min_{h\in H,\nu\in\RR}
\sum_{i=1}^n l(y_i,h(x_i)+\nu) + \frac{\lambda}{2}\|h\|_H^2,
\label{eq:kernel}
\end{equation}  
where the pairs $(x_i,y_i)$ in $X\times \{-1,1\}$, $i=1\dots n$ are training samples, and $l: \{-1,1\}\times \RR\rightarrow\RR^+$ is some arbitrary loss function. 
By definition of a reproducing kernel,
Equation (\ref{eq:kernel}) can be rewritten as
\begin{equation}
\min_{h\in H,\nu\in\RR}
\sum_{i=1}^n l(y_i,\langle \varphi(x_i),h\rangle+\nu) +
\dfrac{\lambda}{2}\|h\|_H^2,
\label{ker:aff}
\end{equation} 
where $\varphi$ is the {\em feature map} over $X$ associated with the
kernel $k$ (which may not admit a known explicit form) and $\langle \cdot, \cdot \rangle$ is the inner product of $H$. We dub problems with the general form of (\ref{ker:aff}) {\em affine}
supervised learning problems since, given some fixed element $h$ of
$H$ and some scalar $\nu$, $\langle h,h'\rangle+\nu$ is an affine function of $h'$,
whose zero set defines an affine hyperplane of $H$ considered itself
as an affine space.

Let $K$ denote the kernel matrix with entries $k_{ij}=\langle\varphi(x_i),
\varphi(x_j)\rangle$ and rows $k_i^T=[k_{i1}, k_{i2},...,k_{in}]$, $i$ in $\{1,\ldots,n\}$.  We assume from now on that $l$ is convex and continuous. Under this assumption, Equation (\ref{ker:aff}) admits an equivalent formulation
\begin{align}
\min_{\alpha\in\RR^{n}, \ \nu\in\RR} \left(\dfrac{1}{n}\sumi l(y_i, k_i^T\alpha+\nu)  +\dfrac{\lambda}{2}\alpha^TK\alpha\right), \label{ker:first}
\end{align}
and any solution $(\alpha^\star,\nu^\star)$ to (\ref{ker:first})
provides
a solution $(h^\star,\nu^\star)$ to (\ref{ker:aff}) with
$h^\star=\sum_{i=1}^n \alpha_i^\star\varphi(x_i)+\nu^\star$. This result follows from the Riesz representation theorem~\cite{SHS01,Wahba90}.

%where $K$ is a $(n+1)\times (n+1)$  matrix and $k_i$ is its $(i+1)$-th column matrix for $i$ in $\{0, 1, ..., n\}$.\\

Assuming our reproducing kernel is semidefinite positive, 
%\footnote{\textit{i.e.} for any $z$ in $\RR^n$, $z^TKz\ge 0$. This is equivalent to all eigenvalues of $K$ being non-negative.} 
$K$ is a semidefinite positive matrix and can be decomposed as $K=BB^T$.
%, with $B$ a rank $r$ upper triangular matrix of size $n\times r$. 
Using this factorization, the kernelized problem can be expressed as 
\begin{equation}
\min_{\beta\in\RR^r,\nu\in\RR} \left( \dfrac{1}{n}\sumi l(y_i, b_i^T\beta+\nu)+\dfrac{\lambda}{2}\|\beta\|^2\right), \label{beta:first}
\end{equation}
where $b_i^T$ denotes the $i$-th row of $B$ and $r$ is the number of columns of $B$. 
If $(\beta^\star, \nu^\star)$ is the solution Equation (\ref{beta:first}), the corresponding vector $\alpha^\star$ (or, more correctly, \textit{a} corresponding vector of dimension $n\geq r$) 
can be computed by $\alpha^\star=P\beta^\star$, where $P$ is the pseudoinverse of $B^T$.

\subsection{Preprocessing of negative samples offline}\label{offline}
Equation (\ref{beta:first}) has the same structure of Equation (\ref{eq:first}) if we were to consider $\theta=0$ and all labels $y_i=-1$.
In that case, all actions described in subsections \ref{slem:intro} and \ref{kernel:review} can be done independently of positive samples. 
They are therefore done offline and their results are stored. In the processing of positive samples, i.e. the online phase, these results are loaded.
The preprocessing phase consists of the calculation of $B$, $\mu = \frac{1}{n}\sum_{i=1}^n b_i^T$ and $A = \frac{1}{n}B^TB-\mu\mu^T +\lambda\mathrm{Id}_r.$

\subsection{Online addition of a positive sample}
\label{subsec:adding}
We now reconsider the exemplar problem, with one positive example $x_0$ and $n$ negative examples $X$. Here $K$ is the $n\times n$ kernel matrix of the negative samples $X$. 
%Let us assume the factor $B$ is given. %The calculation of a kernel Cholesky factorization is solved efficiently by \cite{BaJo05,BaJo02}. 
%We further analyse how to compute this factorization in subsection \ref{offline}.

We denote by $K'$ the augmented kernel matrix obtained by adding the positive samples $x_0$. Such matrix can be written as
\begin{equation}
K' = \begin{bmatrix}
k_{00} & k_0^T\\
k_0 & K
\end{bmatrix},
\end{equation}
where $k_{00}=\langle \varphi(x_0),\varphi(x_0)\rangle$ is a scalar and $k_0= [\langle \varphi(x_0),\varphi(x_i)\rangle]_{1\le i\le n}$ is a vector in $\RR^n$. 
The following lemma show how the factorization of $K'$ can be derived from the factorization of its sub-matrix $K$ and the solution of a $n\times n$ linear system.

\begin{lemma} The augmented kernel matrix $K'$ can be factorized as $K'= B'B'^T$ with
%\begin{align}
%K'&= B'B'^T,\quad\text{where}\quad
%B'=\begin{bmatrix}
%u & v^T\\0 & B
%\end{bmatrix}\\
%v&=B^\dagger k_0,\,\, u=\sqrt{k_{00}-||v||^2}.\label{eq:lemma1}
%\end{align}
\begin{equation}
B'=\begin{bmatrix}
u & v^T\\0 & B
\end{bmatrix},~
v = B^\dagger k_0,~ u=\sqrt{k_{00}-||v||^2}.
\label{eq:lemma1}
\end{equation}
\end{lemma}\label{lemma1}
\begin{proof}
For $B'$ defined by (\ref{eq:lemma1}), we have that
\begin{equation}
B'B'^T = 
\begin{bmatrix} u^2+\|v\|^2 & v^TB^T\\ 
Bv& BB^T\end{bmatrix} 
=\begin{bmatrix} k_{00} & v^TB^T\\ 
Bv& K\end{bmatrix} .
\end{equation}
Since $K'$ is positive semidefinite, $k_0$ must lie in the column space $\mathcal{B}$ of $B$.
%\footnote{
Indeed, if we suppose $k_0$ does not belong to $\mathcal{B}$, then it can be decomposed uniquely as $k_0=s+t$, $s\in\mathcal{B}$ and  $t\in\mathcal{B}^\perp$, with $t\ne 0$. In one hand, $K'$ being semidefinite positive implies that $[1, -at^T]K'[1; -at]=k_{00}-2a\|t\|^2\ge 0$ for all real value $a$. In the other hand, for $a$ large enough, $k_{00}-a\|t\|^2\le 0$, which is a contradiction.
%} 
Hence $v=B^\dagger k_0$ is an exact solution of $Bv=k_0$. The fact that $k_{00}-\|v\|^2$ is non-negative comes from the fact that the Schur complement $K-k_0 k_0^T / k_{00}$ of $k_{00}$ in $K'$ is itself positive semidefinite.
%\footnote{
Indeed, since the matrix $k_{00}K-k_0k_0^T=B(k_{00}\mathrm{Id}_r-vv^T$ is also positive semidefinite. Thus $v^T(k_{00}\mathrm{Id}_r-vv^T)v = \|v\|^2(k_{00}-\|v\|^2)\ge 0$. \qedhere
%}
\end{proof}

This lemma allows us to add a positive sample to Equation (\ref{beta:first}).
%$\beta^\star(x_0,X), \nu^\star(x_0,X) = \underset{\beta\in\RR^{r+1}, \nu\in\RR}{argmin}J'(\beta, \nu)$, for
With a positive exemplar, Equation \ref{beta:first} now reads
\begin{align}
 \theta\ l(1, b_0'^T\beta+\nu) +\dfrac{1}{n}\sumi l(-1,b_i'^T\beta+\nu)
+\dfrac{\lambda}{2}\|\beta\|^2,\label{beta:final}
\end{align}
with $b_i'^T$ being the $(i+1)$-th row of $B'$, $i$ in $\{0,1,...,n\}$. In particular, $b_0'=[u; \ v]$ and, for $i>0$, $b_i'=[0; \ b_i]$. Solution $(\beta^\star,\nu^\star)$ in $\RR^{r+1}\times\RR$ can be computed just as before by Equation (\ref{omega:solution}), replacing $x_0$ by $b_0'$, $\mu$ by $\mu' = \frac{1}{n}\sum_{i=1}^n b_i'$ and $X$ by the $(r+1)\times n$ matrix $Q$ of columns $b_1', b_2',...,b_n'$. $\alpha^\star$ is now calculated as
$\alpha^\star=P'\beta^\star$, where $P' = [u^{-1} \ 0^T; -u^{-1}Pv \ P]$ is the pseudoinverse of $B'^T$. This can be expressed by the linear system

\begin{equation}
\begin{bmatrix} \alpha_0^\star \\ \hat{\alpha}^\star \end{bmatrix} = \begin{bmatrix} \frac{1}{u} & 0^T \\-\frac{1}{u}Pv & P  \end{bmatrix} \begin{bmatrix}\beta_0^\star \\ \hat{\beta}^\star \end{bmatrix}.\label{alpha:to:beta}
\end{equation}

\subsection{Similarity score}\label{simi_score}
Once the optimal parameters $(\beta, \nu)$ from (\ref{beta:final}) and the coordinates $u$, $v$ of $b_0'$ from (\ref{eq:lemma1}) have been found\footnote{We drop the ``$\star$'' in this subsection to avoid cluttering the notation.}, they can be used directly for measuring similarity between matching images.

Indeed, suppose two image representations $x_0$ and $x_0'$ are given and we wish to calculate the similarity score between its SLEM representations $h$ and $h'$, denoted by $s(h,h')$. $h'$ can be written as $h'=\alpha_0'\varphi(x_0')+\sum_{i=1}^n \alpha_i'\varphi (x_i)+\nu'$. Using Equation \ref{alpha:to:beta} and ignoring bias $\nu$ and $\nu'$ which have empirically no influence, $s(h,h')$ is given by:
\begin{equation}
\begin{split}
s(h,h') & = \langle h, h'\rangle \\
		& = \hat{\alpha}^{T} K\hat{\alpha}'+\alpha_0k(X, x_0)^T\hat{\alpha}'+\alpha_0'k(X, x_0')^T\hat{\alpha} \\
		& \ \ \ \ \ +\alpha_0\alpha_0'k(x_0,x_0')\\
		& = \hat{\beta}^T\hat{\beta}'+\frac{\beta_0\beta_0'}{uu'}(k(x_0,x_0')-v^Tv').
\end{split}
\end{equation}

We need to store $x_0$, $\hat{\beta}$ and $v$ to calculate its matching score to whichever other image. Each image depends therefore on a vector of $p+2r$ dimensions.

%%% Local Variables:
%%% TeX-master: "main_eccv"
%%% End: